Analyzing the Impact of Input and Label Modifications on CIFAR-10 Classification

### 2025-05-16 PaperForReview.tex
Added Github URL

%%%%%%%%% ABSTRACT
\begin{abstract}
    The performance of deep learning models is intrinsically tied to the quality and nature of training data. Real-world datasets often suffer from imperfections such as noisy labels or distorted inputs, while augmentations intentionally modify data to improve robustness. This paper investigates the effects of systematic input and label modifications on a standard image classification task using the CIFAR-10 dataset and a ResNet-18 model. We evaluate four distinct scenarios: (1) a baseline with unchanged data, (2) completely randomized labels per sample, (3) 20\% symmetric label noise, and (4) strong input image perturbations. Our findings reveal that while random labels lead to a catastrophic failure in learning (Cohen's $\kappa \approx -0.01$), the model exhibits resilience to 20\% label noise, achieving 64.5\% Top-1 accuracy (vs. 78.3\% baseline). Strikingly, strong input perturbations applied during training, effectively introducing significant data augmentation absent in our unaugmented baseline, act as a powerful regularizer, significantly improving generalization and yielding the highest test accuracy of 85.3\% and a Cohen's Kappa of 0.837. Learning curve analysis further elucidates these effects, showing reduced overfitting with input perturbations. These results underscore the complex interplay between data quality, augmentation strategies, and model robustness, offering insights for developing more resilient machine learning systems.
\end{abstract}

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
    The performance of deep learning models is intrinsically tied to the quality and nature of training data. Real-world datasets often suffer from imperfections such as noisy labels or distorted inputs, while augmentations intentionally modify data to improve robustness. This paper investigates the effects of systematic input and label modifications on a standard image classification task using the CIFAR-10 dataset and a ResNet-18 model. We evaluate four distinct scenarios: (1) a baseline with unchanged data, (2) completely randomized labels per sample, (3) 20\% symmetric label noise, and (4) strong input image perturbations. Our findings reveal that while random labels lead to a catastrophic failure in learning (Cohen's $\kappa \approx -0.01$), the model exhibits resilience to 20\% label noise, achieving 64.5\% Top-1 accuracy (vs. 78.3\% baseline). Strikingly, strong input perturbations applied during training, effectively introducing significant data augmentation absent in our unaugmented baseline, significantly improving generalization and yielding the highest test accuracy of 85.3\% and a Cohen's $\kappa$ of 0.837. These results underscore the complex interplay between data quality, augmentation strategies, and model robustness, offering insights for developing more resilient machine learning systems. The code for this project is available at \url{https://github.com/Rick-McCoy/AI-4190.408-mini-project}.
\end{abstract}
